Provisional Title

### 2021-05-10 draft/draft.tex
starting results, cleaning repo

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\addto{\captionsenglish}{\renewcommand{\bibname}{References}}
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Figure}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions
\hypersetup{citecolor={blue}}
\definecolor{lightgray}{gray}{0.9}
\newcommand{\figref}[2][]{%
  \hyperref[{#2}]{%
    Figure~\ref*{#2}%
    \ifx\\#1\\%
    \else
      #1%
    \fi
  }%
}
\newcommand{\tableref}[1]{%
  \hyperref[{#1}]{%
   Table~\ref*{#1}%
  }%
}
\newcommand{\methodsref}[1]{%
  \hyperref[{#1}]{%
   Methods:~\nameref*{#1}%
  }%
}

@@

\linenumbers

@@

This is an example cite \cite{Luo2020,Vass2021}.
This is how you refer to \figref{fig1}.
This is how you refer to \figref[a]{fig1}.
This is how you refer to \tableref{fig1}.
This is how you refer to \figref{figS1}.
This is how you refer to the section \nameref{methods:sim} of the \nameref{methods}.
This is a cleaner way to refer to the \methodsref{methods:sim}.

@@

\vspace{1.5cm}

Uncovering the rules of microbial community invasions \cite{Vila2019}

@@

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=15cm,keepaspectratio]{figs/fig2.pdf}
\caption{\textbf{Co-selection in microbial community coalescence.}
\textbf{a.}~Coalescence outcomes are quantified by the relative Bray-Curtis similarity
($Q$) between the coalesced and invasive communities (\methodsref{methods:metrics}).
These outcomes are predicted by the pairwise competition between the invasive and
resident dominant
species ($R^2=0.15$ for glutamine and $R^2=0.57$ for citrate). This is consistent with a scenario of
top-down positive co-selection where dominants recruit their cohorts for the final
coalesced community. Two biological replicates per experiment are plotted individually.
\textbf{b.}~We represent the frequency reached by
the invasive dominant species when they invade the resident communities on their own versus
when they are in the company of their cohort.
Three scenarios are possible: green and red
shaded areas represent limit cases of positive (recruitment) or negative (antagonism)
bottom-up co-selection, gray area corresponds to situations where invasive dominant
species can invade with equal success regardless of the presence of their cohorts.
Data shows that positive co-selection is common, whereas antagonistic
co-selection is rare in our experiments.
Two biological replicates per experiment are plotted individually.
\textbf{c-d.} Simulations of community coalescence with a consumer-resource model
are able to capture these trends.}
\label{fig2}
\end{figure}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=15cm,keepaspectratio]{figs/figS1.pdf}
\caption{\textbf{Community compositions after seven additional transfers without coalescence.}
Each color of the pie plots corresponds to a different exact sequence variant
(\methodsref{methods:sequencing}).
Replicate 2 of community 1 from glutamine,
as well as replicate 2 of community 8 from citrate (highlighted) were removed based on their
dissimilarity to the other two replicates (details in code for data analysis, see
\nameref{datacode}).
Communities clustered in dashed boxes shared the same dominant species as revealed by
sequencing data.
For communities enclosed in red boxes, sequencing data showed that the species isolated by plating
was not detectable in the community after seven additional transfers (i.e. the dominant was
incorrectly identified) and were therefore excluded from downstream analyses.}
\label{figS1}
\end{figure}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=12.5cm,keepaspectratio]{figs/figS2.pdf}
\caption{\textbf{Alternative metrics of community distance.}
Quantifying coalescence outcomes using different metrics of community similarity
(\methodsref{methods:metrics}) gives similar results to those shown in \figref[a]{fig2}.
Metrics that account for the relative species abundances (\textbf{a.} Jensen-Shannon similarity or
\figref[a]{fig2} Bray-Curtis similarity) yield higher correlations than less quantitative metrics that only
account for species presence/absence (\textbf{b.} Jaccard similarity or \textbf{c.} Fraction of endemic
invasive species persisting in the coalesced community).}
\label{figS2}
\end{figure}

@@

>>>>

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\addto{\captionsenglish}{\renewcommand{\bibname}{References}}
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Figure}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions
\hypersetup{citecolor={blue}}
\definecolor{lightgray}{gray}{0.9}
\newcommand{\figref}[2][]{%
  \hyperref[{#2}]{%
    Figure~\ref*{#2}%
    \ifx\\#1\\%
    \else
      #1%
    \fi
  }%
}
\newcommand{\tableref}[1]{%
  \hyperref[{#1}]{%
   Table~\ref*{#1}%
  }%
}
\newcommand{\methodsref}[1]{%
  \hyperref[{methods:#1}]{%
   Methods:~\nameref*{methods:#1}%
  }%
}

@@

%\linenumbers

@@

@@

In previous work, we have shown that a large amount of soil and plant microbiomes can be cultured
\textit{ex situ} in synthetic minimal environments with a single supplied limiting resource under serial
growth-dilution cycles \cite{Goldford2018} (\figref[a-b]{fig1}). Under these conditions, environmental
microbiomes spontaneously re-assemble into complex multi-species communities sustained by dense
cross-feeding facilitation networks \cite{Goldford2018}. In addition,
and just like in natural assemblies,
species abundance distributions
in these communities are generally long-tailed and uneven (\figref[d]{fig1} and \figref{figS1}),
with the dominant (most abundant) species typically comprising most of the biomass ($\text{median}=46\%$,
\figref{figS1}). Because these communities are easy to manipulate and grow in high throughput, and
are largely made up by culturable members, they represent good test cases to investigate ecological
co-selection during community coalescence. Here we focus on the dominants and ask whether they can
co-select or be co-selected by the sub-dominant species in their communities (henceforth referred to as
their \textit{cohorts}, \figref[c]{fig1}).
Our results confirm that positive top-down co-selection is common, but its effects are weak.
In contrast, bottom-up co-selection can be very strong, and positive co-selection is far more common
than negative co-selection.
We then turn to a microbial consumer-resource model (microCRM)
\cite{Goldford2018,Marsland2019,Marsland2020}
that is able to capture the dynamics of microbial communities dominated by metabolic interactions,
as is the case for the ones assembled in our experimental conditions \cite{Goldford2018}.
We show that the empirically observed trends in ecological co-selection are reproduced
with minimal model assumptions, and that
tuning the complexity of the metabolic interactions in our \textit{in silico} communities can modulate
the recurrence of top-down or bottom-up co-selection.
Our findings indicate that collective interactions play an important role at dictating community structure
during coalescence.

@@

We collected eight natural microbiomes from different soil and plant environmental samples
(\figref[a]{fig1})
and used them to inoculate our synthetic communities, which were stabilized 
in serial batch-culture biorreactors for 84 generations in synthetic minimal media containing either
glutamine or citrate as the only supplied carbon source
(\figref[b]{fig1}, \methodsref{community-assembly}).
We isolated the dominant species of every community
(\methodsref{dominants})
and identified them by Sanger-sequencing their 16S rRNA gene
(\methodsref{sequencing}), which correctly matched the dominant
Exact Sequence Variant (ESV) \cite{Callahan2016,Callahan2017} found through
community-level 16S illumina sequencing
(\figref{figS1}).
These dominants remained at high frequency after seven additional transfers
with the exception of two of the citrate communities and one of the glutamine
communities (where the dominants were presumably a transiently dominating species)
that were excluded from further analysis
(\figref{figS1}).
Similarly, pairs of communities where the dominants shared a same 16S sequence
and had similar colony morphology were excluded
(\figref{figS1}).

\subsubsection*{Top-dow co-selection is weak}

If communities being coalesced were highly cohesive from the top-down,
the dominant species would co-select the rarer members of its community during coalescence
(\figref[e]{fig1}, left panels).
In this scenario, we would expect the outcome of community coalescence to be predicted by
which of the two dominants is most competitive in pairwise competition.
Coalescence outcomes can be quantified by the similarity between the coalesced and the
invasive communities
(\methodsref{metrics}).
To test this hypothesis, we performed all pairwise competitions between dominant species
in glutamine and citrate environments by mixing them 1:1 on their native media and propagating
the cultures for seven serial transfers, roughly 42 generations
(\methodsref{competitions}).
We then performed all possible pairwise community coalescence experiments by mixing
equal volumes of the communities and propagating the resulting cultures for seven extra
transfers
(\figref[f]{fig1}).
The frequencies of all species in both community-community and dominant-dominant competitions
were determined by 16S illumina sequencing
(\methodsref{sequencing}).
We found that the pairwise competitive ability of an invasive dominant
is only weakly predictive of the performance of the invasive community in coalescence,
as quantified by the relative Bray-Curtis similarity between the coalesced and invasive communities
(\figref[a]{fig2}, $R^2=0.15$, $p<$ for glutamine and $R^2=0.15$, $p<$ for citrate).
Alternative quantifications of community distance also yield weak effects,
and even more so when the metric used
accounts only for the presence/absence of specific species and not for their relative abundance in
the communities
(\figref{figS2}).

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=15cm,keepaspectratio]{figs/fig2.pdf}
\caption{\textbf{Co-selection in microbial community coalescence.}
\textbf{a.}~Coalescence outcomes are quantified by the relative Bray-Curtis similarity
($Q$) between the coalesced and invasive communities (\methodsref{metrics}).
These outcomes are predicted by the pairwise competition between the invasive and
resident dominant
species ($R^2=0.15$ for glutamine and $R^2=0.57$ for citrate). This is consistent with a scenario of
top-down positive co-selection where dominants recruit their cohorts for the final
coalesced community. Two biological replicates per experiment are plotted individually.
\textbf{b.}~We represent the frequency reached by
the invasive dominant species when they invade the resident communities on their own versus
when they are in the company of their cohort.
Three scenarios are possible: green and red
shaded areas represent limit cases of positive (recruitment) or negative (antagonism)
bottom-up co-selection, gray area corresponds to situations where invasive dominant
species can invade with equal success regardless of the presence of their cohorts.
Data shows that positive co-selection is common, whereas antagonistic
co-selection is rare in our experiments.
Two biological replicates per experiment are plotted individually.
\textbf{c-d.} Simulations of community coalescence with a consumer-resource model
are able to capture these trends.}
\label{fig2}
\end{figure}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=15cm,keepaspectratio]{figs/figS1.pdf}
\caption{\textbf{Community compositions after seven additional transfers without coalescence.}
Each color of the pie plots corresponds to a different exact sequence variant
(\methodsref{sequencing}).
Replicate 2 of community 1 from glutamine,
as well as replicate 2 of community 8 from citrate (highlighted) were removed based on their
dissimilarity to the other two replicates (details in code for data analysis, see
\nameref{datacode}).
Communities clustered in dashed boxes shared the same dominant species as revealed by
sequencing data.
For communities enclosed in red boxes, sequencing data showed that the species isolated by plating
was not detectable in the community after seven additional transfers (i.e. the dominant was
incorrectly identified) and were therefore excluded from downstream analyses.}
\label{figS1}
\end{figure}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=12.5cm,keepaspectratio]{figs/figS2.pdf}
\caption{\textbf{Alternative metrics of community distance.}
Quantifying coalescence outcomes using different metrics of community similarity
(\methodsref{metrics}) gives similar results to those shown in \figref[a]{fig2}.
Metrics that account for the relative species abundances (\textbf{a.} Jensen-Shannon similarity or
\figref[a]{fig2} Bray-Curtis similarity) yield higher correlations than less quantitative metrics that only
account for species presence/absence (\textbf{b.} Jaccard similarity or \textbf{c.} Fraction of endemic
invasive species persisting in the coalesced community).}
\label{figS2}
\end{figure}

@@

This is an example cite \cite{Luo2020,Vass2021}.
This is how you refer to \figref{fig1}.
This is how you refer to \figref[a]{fig1}.
This is how you refer to \tableref{fig1}.
This is how you refer to \figref{figS1}.
This is how you refer to the section \nameref{methods:sim} of the \nameref{methods}.
This is a cleaner way to refer to the \methodsref{sim}.
